---
title: "RSP analysis"
author: "Katie L. Duggan DiDominic"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Load packages
```{r}
#install.packages("tidyverse")
library(tidyverse)
#install.packages("NatParksPalettes")
library(NatParksPalettes)
#install.packages("ggpmisc")
library(ggpmisc)
library(ggthemes)
library(lmtest)
```

## Load data
```{r}
rspdata <- read_csv(file = "/Users/u4rrnkld/OneDrive - US Army Corps of Engineers/Documents/RSPtoCBR_correlation/data/rspdata.csv")

cbrdata <- read_csv(file = "/Users/u4rrnkld/OneDrive - US Army Corps of Engineers/Documents/RSPtoCBR_correlation/data/cbrdata.csv")

meancbrdata <- read_csv(file = "/Users/u4rrnkld/OneDrive - US Army Corps of Engineers/Documents/RSPtoCBR_correlation/data/meancbrdata.csv") %>% 
  select(c(station_test, hours_avg, CBR))
```

## Analyze

### Spaghetti plots for all data
```{r}
# Create a small dataframe for the C-17 segments
c17_segment_data <- data.frame(
  x = c(99, 65),
  xend = c(99, 65),
  y = c(0, 250),
  yend = c(250, 400),
  segment_label = "Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs")  # label that will show in the legend

c17_segment_data$sample_id <- NA

# Spaghetti Plot
colorblind_colors2 <- c("CBR1" = "#E69F00", "CBR2" = "#56B4E9", "CBR3" = "#009E73", "CBR5" = "#F0E442", "CBR4" = "#CC79A7") 

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>% 
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = sample_id, color = CBR_id)) +
  geom_path(linewidth = 0.75) +
  #geom_vline(data = station_test_mean_unfiltered, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  geom_segment(data = c17_segment_data,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "red3", 
               inherit.aes = FALSE,
               linewidth = 0.75) +
  #scale_linetype_manual(name = "",
                        #values = c("Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs" = "solid", "Mean RSP" = "dashed")) +
  scale_color_manual(values = colorblind_colors2) +
  facet_grid(station~test, switch = "x") +
  scale_y_continuous(trans = "reverse") +
  scale_x_continuous(position = "top") +
  xlab("RSP Index") +
  ylab("Depth from Surface (mm)") +
  labs(color = "Corresponding\nCBR Test\nLocation") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15))

# Average spaghetti plot for each time point (test #)
cbr_avg <- rspdata %>% 
  select(c(sample_id, depth_from_surface, rsp_index)) %>% 
  pivot_wider(names_from = sample_id, values_from = rsp_index) %>% 
  arrange(depth_from_surface) %>% 
  mutate(STA1_test0 = round(rowMeans(across(starts_with("STA1_test0")), na.rm=TRUE))) %>% 
  mutate(STA2_test0 = round(rowMeans(across(starts_with("STA2_test0")), na.rm=TRUE))) %>% 
  mutate(STA1_test1 = round(rowMeans(across(starts_with("STA1_test1")), na.rm=TRUE))) %>% 
  mutate(STA2_test1 = round(rowMeans(across(starts_with("STA2_test1")), na.rm=TRUE))) %>%
  mutate(STA1_test2 = round(rowMeans(across(starts_with("STA1_test2")), na.rm=TRUE))) %>% 
  mutate(STA2_test2 = round(rowMeans(across(starts_with("STA2_test2")), na.rm=TRUE))) %>%
  mutate(STA1_test3 = round(rowMeans(across(starts_with("STA1_test3")), na.rm=TRUE))) %>% 
  mutate(STA2_test3 = round(rowMeans(across(starts_with("STA2_test3")), na.rm=TRUE))) %>%
  mutate(STA1_test4 = round(rowMeans(across(starts_with("STA1_test4")), na.rm=TRUE))) %>% 
  mutate(STA2_test4 = round(rowMeans(across(starts_with("STA2_test4")), na.rm=TRUE))) %>%
  select(c(STA1_test0, STA2_test0, STA1_test1, STA2_test1, STA1_test2, STA2_test2, STA1_test3, STA2_test3, STA1_test4, STA2_test4, depth_from_surface)) %>% 
  pivot_longer(cols = c(STA1_test0, STA2_test0, STA1_test1, STA2_test1, STA1_test2, STA2_test2, STA1_test3, STA2_test3, STA1_test4, STA2_test4), names_to = "station_test", values_to = "rsp_index") %>% 
  mutate(station = substr(station_test, 1, 4)) %>% 
  mutate(test = substr(station_test, 6, 10))

cbr_avg %>% 
  mutate(test = ifelse(test == "test0", "Test 0", test)) %>% 
  mutate(test = ifelse(test == "test1", "Test 1", test)) %>% 
  mutate(test = ifelse(test == "test2", "Test 2", test)) %>% 
  mutate(test = ifelse(test == "test3", "Test 3", test)) %>% 
  mutate(test = ifelse(test == "test4", "Test 4", test)) %>% 
  mutate(test = ifelse(test == "test5", "Test 5", test)) %>% 
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = station_test, color = test)) +
  geom_path(linewidth = 0.75) +
  #scale_color_manual(values = test_colors) +
  scale_color_manual(values=natparks.pals("Yellowstone", 5)) +
  facet_grid(~station, switch = "x") +
  scale_y_continuous(trans = "reverse") +
  scale_x_continuous(position = "top") +
  xlab("Average RSP Index") +
  ylab("Depth from Surface (mm)") +
  labs(color = "Corresponding\nCBR Timepoint") +
  theme_bw()
```

### Stepwise versions (NOT USED)
```{r}
# STEPwise version of previous plot - (NOT USED SO FAR)
rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>% 
  ggplot(aes(x = depth_from_surface, y = rsp_index, group = sample_id, color = CBR_id)) +
  geom_step(linewidth = 0.75, direction = "vh") +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_grid(station~test, switch = "x") +
  scale_x_continuous(trans = "reverse") +
  scale_y_continuous(position = "right") +
  coord_flip() +
  xlab("Depth from Surface (mm)") +
  ylab("RSP Index") +
  labs(color = "Corresponding\nCBR") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12))

### Each CBR and each test (colored by station) - (NOT USED SO FAR)
colors <- natparks.pals("Yellowstone", 5)
selected_colors <- colors[c(1, 3, 4)]

rspdata %>% ## stepwise station vs station grid
  filter(station == "STA1") %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>% 
  ggplot(aes(x = depth_from_surface, y = rsp_index, group = sample_id, color = rep)) +
  geom_step(linewidth = 0.75, direction = "vh") +
  scale_color_manual(values=selected_colors) +
  facet_grid(CBR_id~test, switch = "x") +
  scale_x_continuous(trans = "reverse") +
  scale_y_continuous(position = "right") +
  coord_flip() +
  xlab("Depth from Surface (mm)") +
  ylab("RSP Index") +
  labs(color = "Corresponding\nCBR") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12))
```

### Spaghetti plots for each test (facet by station)
```{r}
# Test 0
test0_means <- station_test_mean_unfiltered %>% 
  filter(test == "Test 0")

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>%
  filter(test == "Test 0") %>% 
  #View()
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = sample_id, color = CBR_id)) +
  geom_path(linewidth = 0.75) +
  #geom_vline(data = test0_means, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  geom_segment(data = c17_segment_data,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "darkred", 
               inherit.aes = FALSE,
               linewidth = 0.75) +
  #scale_linetype_manual(name = "",
                        #values = c("Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs" = "solid", "Mean RSP" = "dashed")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_wrap(~station, strip.position = "bottom") +
  scale_y_continuous(trans = "reverse", name = "Penetration Depth (mm)", 
                     sec.axis = sec_axis(~ . / 25.4, name = "Penetration Depth (in)")) +
  scale_x_continuous(position = "top", limit = c(0, 200)) +
  xlab("RSP Index") +
  labs(color = "Corresponding\nCBR Test\nLocation") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15),
        axis.title.y.right = element_text(angle = 90))
        #panel.grid.major = element_line(color = "grey80"))

# Test 1
test1_means <- station_test_mean_unfiltered %>% 
  filter(test == "Test 1")

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>%
  filter(test == "Test 1") %>% 
  #View()
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = sample_id, color = CBR_id)) +
  geom_path(linewidth = 0.75) +
  #geom_vline(data = test0_means, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  geom_segment(data = c17_segment_data,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "darkred", 
               inherit.aes = FALSE,
               linewidth = 0.75) +
  #scale_linetype_manual(name = "",
                        #values = c("Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs" = "solid", "Mean RSP" = "dashed")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_wrap(~station, strip.position = "bottom") +
  scale_y_continuous(trans = "reverse", name = "Penetration Depth (mm)", 
                     sec.axis = sec_axis(~ . / 25.4, name = "Penetration Depth (in)")) +
  scale_x_continuous(position = "top", limit = c(0, 200)) +
  xlab("RSP Index") +
  labs(color = "Corresponding\nCBR Test\nLocation") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15),
        axis.title.y.right = element_text(angle = 90))
        #panel.grid.major = element_line(color = "grey80"))

# Test 2
test2_means <- station_test_mean_unfiltered %>% 
  filter(test == "Test 2")

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>%
  filter(test == "Test 2") %>% 
  #View()
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = sample_id, color = CBR_id)) +
  geom_path(linewidth = 0.75) +
  #geom_vline(data = test0_means, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  geom_segment(data = c17_segment_data,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "darkred", 
               inherit.aes = FALSE,
               linewidth = 0.75) +
  #scale_linetype_manual(name = "",
                        #values = c("Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs" = "solid", "Mean RSP" = "dashed")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_wrap(~station, strip.position = "bottom") +
  scale_y_continuous(trans = "reverse", name = "Penetration Depth (mm)", 
                     sec.axis = sec_axis(~ . / 25.4, name = "Penetration Depth (in)")) +
  scale_x_continuous(position = "top", limit = c(0, 200)) +
  xlab("RSP Index") +
  labs(color = "Corresponding\nCBR Test\nLocation") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15),
        axis.title.y.right = element_text(angle = 90))
        #panel.grid.major = element_line(color = "grey80"))

# Test 3
test3_means <- station_test_mean_unfiltered %>% 
  filter(test == "Test 3")

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>%
  filter(test == "Test 3") %>% 
  #View()
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = sample_id, color = CBR_id)) +
  geom_path(linewidth = 0.75) +
  #geom_vline(data = test0_means, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  geom_segment(data = c17_segment_data,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "darkred", 
               inherit.aes = FALSE,
               linewidth = 0.75) +
  #scale_linetype_manual(name = "",
                        #values = c("Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs" = "solid", "Mean RSP" = "dashed")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_wrap(~station, strip.position = "bottom") +
  scale_y_continuous(trans = "reverse", name = "Penetration Depth (mm)", 
                     sec.axis = sec_axis(~ . / 25.4, name = "Penetration Depth (in)")) +
  scale_x_continuous(position = "top", limit = c(0, 200)) +
  xlab("RSP Index") +
  labs(color = "Corresponding\nCBR Test\nLocation") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15),
        axis.title.y.right = element_text(angle = 90))
        #panel.grid.major = element_line(color = "grey80"))

# Test 4
test4_means <- station_test_mean_unfiltered %>% 
  filter(test == "Test 4")

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>%
  filter(test == "Test 4") %>% 
  #View()
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = sample_id, color = CBR_id)) +
  geom_path(linewidth = 0.75) +
  #geom_vline(data = test0_means, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  geom_segment(data = c17_segment_data,
               aes(x = x, xend = xend, y = y, yend = yend),
               color = "darkred", 
               inherit.aes = FALSE,
               linewidth = 0.75) +
  #scale_linetype_manual(name = "",
                        #values = c("Minimum RSP\nRequired for\nAircraft Load\nof 500,000 lbs" = "solid", "Mean RSP" = "dashed")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_wrap(~station, strip.position = "bottom") +
  scale_y_continuous(trans = "reverse", name = "Penetration Depth (mm)", 
                     sec.axis = sec_axis(~ . / 25.4, name = "Penetration Depth (in)")) +
  scale_x_continuous(position = "top", limit = c(0, 200)) +
  xlab("RSP Index") +
  labs(color = "Corresponding\nCBR Test\nLocation") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15),
        axis.title.y.right = element_text(angle = 90))
        #panel.grid.major = element_line(color = "grey80"))
```

### Mean RSP for Correlation
```{r}
# Calculate mean lines
station_test_mean_unfiltered <- rspdata %>%
  #group_by(cbr_id) %>%
  group_by(station_test) %>%
  summarize(mean_rsp = mean(rsp_index)) %>% 
  mutate(station = str_sub(station_test, 1, 4)) %>% 
  mutate(test = str_sub(station_test, 6, 10)) %>%
  mutate(test = ifelse(test == "test0", "Test 0", test)) %>% 
  mutate(test = ifelse(test == "test1", "Test 1", test)) %>% 
  mutate(test = ifelse(test == "test2", "Test 2", test)) %>% 
  mutate(test = ifelse(test == "test3", "Test 3", test)) %>% 
  mutate(test = ifelse(test == "test4", "Test 4", test)) %>% 
  mutate(test = ifelse(test == "test5", "Test 5", test))

rspdata %>% 
  mutate(CBR_rep = paste(CBR_id, rep)) %>% 
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = CBR_id, color = CBR_id)) +
  #geom_path(linewidth = 0.75) +
  geom_point() +
  geom_vline(data = station_test_mean_unfiltered, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  #scale_linetype_manual(name = "Minimum Required \nRSP for Aircraft Load",
                        #values = c("500,000 lbs" = "solid")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_grid(station~test, switch = "x") +
  scale_y_continuous(trans = "reverse") +
  scale_x_continuous(position = "top") +
  xlab("RSP Index") +
  ylab("Depth from Surface (mm)") +
  labs(color = "Corresponding\nCBR") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12))

####### plot of mean RSPs (all data)
mean_rsp_cbr <- full_join(meancbrdata, station_test_mean_unfiltered, by = "station_test")

visual_mean_rsp <- mean_rsp_cbr %>% 
  select(-mean_rsp) %>% 
  mutate(visual_mean_rsp = c(25, 29.5, 34, 30, 32, 30, 41.7, 45.3, 48.3, 54))

group_colors <- c("STA1" = "darkgreen", "STA2" = "deepskyblue")

mean_rsp_cbr %>% 
  ggplot(aes(x = hours_avg, y = mean_rsp, group = station, color = station)) +
  geom_line(linewidth = 1.25) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dotted") +
  #stat_poly_eq(
    #aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    #formula = y ~ x,
    #parse = TRUE) +
  scale_color_manual(values = group_colors) +
  scale_x_continuous(limits = c(0,35), breaks = seq(0, 35, by = 5)) +
  xlab("Time Duration (hr)") +
  ylab("Average RSP across depth (0-400mm)") +
  labs(color = "Test Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15))
## stat_poly_eq gives the following equations (add them manually in ppt to match Margie's plot)
#### green: y = 25.8 + 0.21x (R2 = 0.81)
#### blue: y = 33 + 0.57x (R2 = 0.89)

visual_mean_rsp %>% #### Use this one in TR/manuscript since it uses the "visual average" values
  ggplot(aes(x = hours_avg, y = visual_mean_rsp, group = station, color = station)) +
  geom_line(linewidth = 1.25) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dotted") +
  #stat_poly_eq(
    #aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    #formula = y ~ x,
    #parse = TRUE) +
  scale_color_manual(values = group_colors) +
  scale_x_continuous(limits = c(0,35), breaks = seq(0, 35, by = 5)) +
  xlab("Time Duration (hr)") +
  ylab("Mean RSP Index") +
  labs(color = "Test Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15))
## stat_poly_eq gives the following equations (add them manually in ppt to match Margie's plot)
#### green: y = 26.7 + 0.199x (R2 = 0.62)
#### blue: y = 33.5 + 0.631x (R2 = 0.84)

### George comment on TR: look into the difference between STA 1 and STA 2
no_test_0 <- visual_mean_rsp %>% 
  filter(!station_test == "STA1_test0")
  
no_test_0 %>% 
  ggplot(aes(x = hours_avg, y = visual_mean_rsp, group = station, color = station)) +
  geom_line(linewidth = 1.25) +
  geom_point(size = 3) +
  geom_smooth(method = "lm", se = FALSE, linetype = "dotted") +
  stat_poly_eq(
    aes(label = paste(..eq.label.., ..rr.label.., sep = "~~~")),
    formula = y ~ x,
    parse = TRUE) +
  scale_color_manual(values = group_colors) +
  scale_x_continuous(limits = c(0,35), breaks = seq(0, 35, by = 5)) +
  xlab("Time Duration (hr)") +
  ylab("Mean RSP Index") +
  labs(color = "Test Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15))
```

#### Correlation with mean RSP from points (10 to 10, CBR to RSP)
```{r}
rsp_means_for_corr <- rspdata %>% #filter for RSPs relating to the CBRs that Margie kept
  filter(!(station == "STA1" & test == "Test 0" & CBR_id == "CBR4")) %>% # remove CBR4 for STA1/test0
  filter(!(station == "STA1" & test == "Test 1" & CBR_id == "CBR3")) %>% # remove CBR3 for STA1/test1
  filter(!(station == "STA1" & test == "Test 1" & CBR_id == "CBR4")) %>% # remove CBR4 for STA1/test1
  filter(!(station == "STA1" & test == "Test 2" & CBR_id == "CBR1")) %>% # remove CBR1 for STA1/test2
  filter(!(station == "STA1" & test == "Test 3" & CBR_id == "CBR3")) %>% # remove CBR3 for STA1/test3
  filter(!(station == "STA2" & test == "Test 2" & CBR_id == "CBR1")) %>% # remove CBR1 for STA2/test2
  filter(between(depth_from_surface, 50, 200)) %>% #keep 50mm through 200mm)
  group_by(station_test) %>%
  summarize(mean_rsp = mean(rsp_index)) %>% 
  mutate(station = str_sub(station_test, 1, 4)) %>% 
  mutate(test = str_sub(station_test, 6, 10)) %>%
  mutate(test = ifelse(test == "test0", "Test 0", test)) %>% 
  mutate(test = ifelse(test == "test1", "Test 1", test)) %>% 
  mutate(test = ifelse(test == "test2", "Test 2", test)) %>% 
  mutate(test = ifelse(test == "test3", "Test 3", test)) %>% 
  mutate(test = ifelse(test == "test4", "Test 4", test)) %>% 
  mutate(test = ifelse(test == "test5", "Test 5", test))

rspdata %>% 
  filter(!(station == "STA1" & test == "Test 0" & CBR_id == "CBR4")) %>% # remove CBR4 for STA1/test0
  filter(!(station == "STA1" & test == "Test 1" & CBR_id == "CBR3")) %>% # remove CBR3 for STA1/test1
  filter(!(station == "STA1" & test == "Test 1" & CBR_id == "CBR4")) %>% # remove CBR4 for STA1/test1
  filter(!(station == "STA1" & test == "Test 2" & CBR_id == "CBR1")) %>% # remove CBR1 for STA1/test2
  filter(!(station == "STA1" & test == "Test 3" & CBR_id == "CBR3")) %>% # remove CBR3 for STA1/test3
  filter(!(station == "STA2" & test == "Test 2" & CBR_id == "CBR1")) %>% # remove CBR1 for STA2/test2
  filter(between(depth_from_surface, 50, 200)) %>% #keep 50mm through 200mm)
  mutate(CBR_rep = paste(CBR_id, rep)) %>% 
  ggplot(aes(x = rsp_index, y = depth_from_surface, group = CBR_id, color = CBR_id)) +
  #geom_path(linewidth = 0.75) +
  geom_point() +
  geom_vline(data = rsp_means_for_corr, aes(xintercept = mean_rsp), color = "black", linetype = "dashed", linewidth = 1) +
  #scale_linetype_manual(name = "Minimum Required \nRSP for Aircraft Load",
                        #values = c("500,000 lbs" = "solid")) +
  scale_color_manual(values = natparks.pals("Yellowstone", 5)) +
  facet_grid(station~test, switch = "x") +
  scale_y_continuous(trans = "reverse") +
  scale_x_continuous(position = "top") +
  xlab("RSP Index") +
  ylab("Depth from Surface (mm)") +
  labs(color = "Corresponding\nCBR") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12))

means_for_corr <- full_join(meancbrdata, rsp_means_for_corr, by = "station_test")

means_for_corr %>% ## only those relating to remaining CBR data
  ggplot(aes(x = hours_avg, y = mean_rsp, group = station, color = station)) +
  geom_line(linewidth = 1) +
  #scale_color_manual(values = selected_colors) +
  xlab("Time since Test 0 (hours)") +
  ylab("Mean RSP across depth (0-400mm)") +
  labs(color = "Test Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 12),
        axis.title = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        strip.text = element_text(size = 12))

write.csv(means_for_corr, "/Users/u4rrnkld/OneDrive - US Army Corps of Engineers/Documents/RSPtoCBR_correlation/data/means_for_corr_2025.csv", row.names = FALSE)
```

## remake CBR and RSP plots for manuscript
#### Station-CBR replicates plot
```{r}
colorblind_colors <- c("STA1-R1" = "#000000", "STA1-R2" = "#E69F00", "STA1-R3" = "#56B4E9", "STA2-R1" = "#009E73", "STA2-R2" = "#F0E442", "STA2-R3" = "#CC79A7") 

cbrdata %>% 
  mutate(replicate = paste0("R", replicate)) %>% 
  mutate(station_CBRrep = paste(station, replicate, sep = "-")) %>% 
  ggplot(aes(x = time_hours, y = CBR, group = station_CBRrep, color = station_CBRrep)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  #scale_color_colorblind() +
  scale_color_manual(values = colorblind_colors) +
  xlab("Time (hr)") +
  ylab("CBR") +
  labs(color = "Station-CBR \nReplicate") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        strip.text = element_text(size = 15))
```

#### CBR for each station with trendlines
```{r}
STA_colors <- c("STA1" = "#CB7223", "STA2" = "#0067A2") 

## CBR trend lines
mean_sd_cbr <- cbrdata %>% 
  group_by(station_test) %>% 
  summarize(CBR_mean = mean(CBR, na.rm = TRUE),
            CBR_sd = sd(CBR, na.rm = TRUE),
            hours_avg = mean(time_hours),
            .groups = "drop")

mean_sd_cbr_combined <- full_join(cbrdata, mean_sd_cbr, by = "station_test")

cbr_trend_eq_1 <- paste0("y = ", 0.152, "x + ", 8.39)
cbr_trend_eq_2 <- paste0("y = ", 0.356, "x + ", 7.43)

mean_sd_cbr_combined %>% 
  mutate(station = str_sub(station_test, 1, 4)) %>%
  ggplot(aes(x = hours_avg, y = CBR_mean, group = station, color = station)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_abline(slope = 0.15221, intercept = 8.38959, linetype = "dashed", linewidth = 1, color = "#CB7223") + #station 1 line
  annotate("text", x = 16, y = 9, label = cbr_trend_eq_1, color = "#CB7223", size = 5.5) + #station 1 eq
  annotate("text", x = 16, y = 8, label = "R^2 == 0.959", parse = TRUE, color = "#CB7223", size = 5.5) + #station 1 r2
  geom_abline(slope = 0.35575, intercept = 7.43043, linetype = "dashed", linewidth = 1, color = "#0067A2") + #station 2 line
  annotate("text", x = 13, y = 15, label = cbr_trend_eq_2, color = "#0067A2", size = 5.5) + #station 2 eq
  annotate("text", x = 13, y = 14, label = "R^2 == 0.954", parse = TRUE, color = "#0067A2", size = 5.5) + #station 2 r2
  #geom_smooth(method = "lm", se = FALSE, linetype = "dotted") +
  geom_errorbar(aes(ymin = CBR-CBR_sd, ymax = CBR+CBR_sd)) +
  scale_color_manual(values = STA_colors) +
  xlab("Time (hr)") +
  ylab("CBR") +
  labs(color = "Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.position = "inside",
        legend.position.inside = c(0.1, 0.875),
        strip.text = element_text(size = 15))

#### per reviewer comment: try other non-linear fits
cbr_sta1 <- mean_sd_cbr %>% 
  filter(str_starts(station_test, "STA1"))

cbr_sta2 <- mean_sd_cbr %>% 
  filter(str_starts(station_test, "STA2"))

# sta1 linear
mod_cbr_sta1_lm = lm(CBR_mean ~ hours_avg, data=cbr_sta1)
summary(mod_cbr_sta1_lm)
raintest(mod_cbr_sta1_lm)

ggplot(mod_cbr_sta1_lm, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Residuals vs. Fitted Values Plot (Simple Method)",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

#sta1 exponential
mod_cbr_sta1_exp = lm(log(CBR_mean) ~ hours_avg, data=cbr_sta1)
summary(mod_cbr_sta1_exp)

#sta1 power
cbr_sta1_pow <- cbr_sta1 %>% 
  mutate(time_hours_adj = ifelse(hours_avg == 0, 0.01, hours_avg)) #change 0s to 0.01 (36 seconds) to avoid log(0) error

mod_cbr_sta1_pow = lm(log(CBR_mean) ~ log(time_hours_adj), data=cbr_sta1_pow)
summary(mod_cbr_sta1_pow)

#sta2 linear
mod_cbr_sta2_lm = lm(CBR_mean ~ hours_avg, data=cbr_sta2)
summary(mod_cbr_sta2_lm)
raintest(mod_cbr_sta2_lm)

ggplot(mod_cbr_sta2_lm, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Residuals vs. Fitted Values Plot (Simple Method)",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

#sta2 exponential
mod_cbr_sta2_exp = lm(log(CBR_mean) ~ hours_avg, data=cbr_sta2)
summary(mod_cbr_sta2_exp)

#sta2 power
cbr_sta2_pow <- cbr_sta2 %>% 
  mutate(time_hours_adj = ifelse(hours_avg == 0, 0.01, hours_avg))

mod_cbr_sta2_pow = lm(log(CBR_mean) ~ log(time_hours_adj), data=cbr_sta2_pow)
summary(mod_cbr_sta2_pow)
```

#### RSP for each station with trendlines
```{r}
## RSP trend lines
rsp_trend_eq_1 <- paste0("y = ", 0.199, "x + ", 26.7)
rsp_trend_eq_2 <- paste0("y = ", 0.631, "x + ", 33.5)

visual_mean_rsp %>% #### Use this one in TR since it used the "visual average" values
  ggplot(aes(x = hours_avg, y = visual_mean_rsp, group = station, color = station)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  geom_abline(slope = 0.19851, intercept = 26.65791, linetype = "dashed", linewidth = 1, color = "#CB7223") + #station 1 line
  annotate("text", x = 20, y = 28.5, label = rsp_trend_eq_1, color = "#CB7223", size = 5.5) + #station 1 eq
  annotate("text", x = 20, y = 27, label = "R^2 == 0.624", parse = TRUE, color = "#CB7223", size = 5.5) + #station 1 r2
  geom_abline(slope = 0.6308, intercept = 33.4522, linetype = "dashed", linewidth = 1, color = "#0067A2") + #station 2 line
  annotate("text", x = 14, y = 47, label = rsp_trend_eq_2, color = "#0067A2", size = 5.5) + #station 2 eq
  annotate("text", x = 14, y = 45.5, label = "R^2 == 0.845", parse = TRUE, color = "#0067A2", size = 5.5) + #station 2 r2
  #geom_smooth(method = "lm", se = FALSE, linetype = "dotted") +
  scale_color_manual(values = STA_colors) +
  scale_x_continuous(limits = c(0, 32), breaks = seq(0, 32, by = 4)) +
  xlab("Time (hr)") +
  ylab("Mean RSP Index") +
  labs(color = "Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.position = "inside",
        legend.position.inside = c(0.1, 0.875),
        strip.text = element_text(size = 15))
## stat_poly_eq gives the following equations (add them manually in ppt to match Margie's plot)
#### green: y = 26.7 + 0.199x (R2 = 0.62)
#### blue: y = 33.5 + 0.631x (R2 = 0.84)

#### per reviewer comment: try other non-linear fits
rsp_sta1 <- visual_mean_rsp %>% 
  filter(station == "STA1")

rsp_sta2 <- visual_mean_rsp %>% 
  filter(station == "STA2")

# sta1 linear
mod_rsp_sta1_lm = lm(visual_mean_rsp ~ hours_avg, data=rsp_sta1)
summary(mod_rsp_sta1_lm)
raintest(mod_rsp_sta1_lm)

ggplot(mod_rsp_sta1_lm, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Residuals vs. Fitted Values Plot (Simple Method)",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

#sta1 exponential
mod_rsp_sta1_exp = lm(log(visual_mean_rsp) ~ hours_avg, data=rsp_sta1)
summary(mod_rsp_sta1_exp)

#sta1 power
rsp_sta1_pow <- rsp_sta1 %>% 
  mutate(time_hours_adj = ifelse(hours_avg == 0, 0.01, hours_avg)) #change 0s to 0.01 (36 seconds) to avoid log(0) error

mod_rsp_sta1_pow = lm(log(visual_mean_rsp) ~ log(time_hours_adj), data=rsp_sta1_pow)
summary(mod_rsp_sta1_pow)

#sta2 linear
mod_rsp_sta2_lm = lm(visual_mean_rsp ~ hours_avg, data=rsp_sta2)
summary(mod_rsp_sta2_lm)
raintest(mod_rsp_sta2_lm)

ggplot(mod_rsp_sta2_lm, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Residuals vs. Fitted Values Plot (Simple Method)",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

#sta2 exponential
mod_rsp_sta2_exp = lm(log(visual_mean_rsp) ~ hours_avg, data=rsp_sta2)
summary(mod_rsp_sta2_exp)

#sta2 power
rsp_sta2_pow <- rsp_sta2 %>% 
  mutate(time_hours_adj = ifelse(hours_avg == 0, 0.01, hours_avg))

mod_rsp_sta2_pow = lm(log(visual_mean_rsp) ~ log(time_hours_adj), data=rsp_sta2_pow)
summary(mod_rsp_sta2_pow)
```

#### Correlation plot (CBR vs RSP)
```{r}
## correlation model
#linear
corr_lm <- lm(CBR ~ mean_rsp, data = means_for_corr)
corr_lm_r2 <- summary(corr_lm)$r.squared

ggplot(corr_lm, aes(x = .fitted, y = .resid)) +
  geom_point() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed", size = 1) +
  labs(title = "Residuals vs. Fitted Values Plot (Simple Method)",
       x = "Fitted Values",
       y = "Residuals") +
  theme_minimal()

#exponential
corr_exp <- lm(log(CBR) ~ mean_rsp, data = means_for_corr)
corr_exp_r2 <- summary(corr_exp)$r.squared

#power law
corr_pow = lm(log(CBR) ~ log(mean_rsp), data=means_for_corr)
corr_pow_r2 <- summary(corr_pow)$r.squared

#model equation:
equation_string <- paste0("y = ", 0.4600, "x - ", 3.76)

## plot
ggplot(means_for_corr, aes(x = mean_rsp, y = CBR)) +
  geom_point(aes(color = station), size = 3) +
  #scale_x_log10() +
  #scale_y_log10() +
  scale_color_manual(values = STA_colors) + #(use group_colors for the TR)
  geom_abline(slope = 0.4600, intercept = -3.7572, linetype = "dashed", linewidth = 1) +
  annotate("text", x = 30, y = 20, label = equation_string, color = "black", size = 5.5) +
  annotate("text", x = 30, y = 18, label = "R^2 == 0.44", parse = TRUE, color = "black", size = 5.5) +
  xlab("RSP Index") +
  ylab("CBR") +
  xlim(0, 45) + #(making the scale the same helps see that the trend line equation is correct)
  ylim(0, 45) +
  labs(color = "Station") +
  theme_bw() +
  theme(axis.text = element_text(size = 15),
        axis.title = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.title = element_text(size = 15),
        legend.position = "inside",
        legend.position.inside = c(0.1, 0.875),
        strip.text = element_text(size = 15))
```




